---
title: "Covid91 vaccine study Final2024"
author: "Mallori Rojas"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "`r Sys.Date()`" 
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
FinalDat <-read.csv(file="finalSumr2024.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- FinalDat  %>% select(-one_of("X"))
FinalData <- FinalData %>%
mutate(across(where(is.character), as.factor))
```

# Introduction


A new preventative vaccine (covid921) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbiage when referencing p-values in conclusions.




## Males


$H_0$: The Covid921 vaccine is not effective in males

$H_A$: The Covid921 vaccine is effective in males



### Methods





Treatment will be Cat
Male will be Cat
helper file we are using is
CAT~CAT

### Descriptive Results

#### Graphical Results


```{r}
males=filter(FinalData,FinalData$sex=="M")
barchartGC(~treatment + infected,data=males)
barchartGC(~infected + treatment,data=males,type="percent")

```



This graph shows that the MALES who got the placebo were more infected than the MALES that were vaccinated, but about 50% that got placebos were OK, and about 50% of MALES with the vaccine were OK.


#### Numerical Results
```{r}
table1 <- xtabs(~treatment + infected,data=males)
rowPerc(table1)
colPerc(table1)
```


This shows that about 74% males who took the placebo got covid19, where about 26% of males who got the vaccines got covid19



#### Inferential Results

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```


The statistical analysis reveals a notable difference in COVID-19 infection rates between the placebo and vaccine groups. Both the chi-square test and Fisher’s Exact Test yielded a p-value of 0.00000000022, indicating a strong reduction in the probability of contracting COVID-19 for those receiving the vaccine. Specifically, individuals receiving the placebo have about 2.82 times higher odds of contracting COVID-19 compared to those who received the vaccine. The 95% confidence interval supports the vaccine’s effectiveness in preventing COVID-19 with a high level of statistical significance.






## Females

$H_0$: The Covid921 vaccine is not effective in females

$H_A$: The Covid921 vaccine is effective in females

### Methods


Female is CAT
Treatment is CAT
Helper File we use
CAT ~ CAT

### Descriptive Results

#### Graphical Results

```{r}
females=filter(FinalData,FinalData$sex=="F")
barchartGC(~treatment + infected,data=females)
barchartGC(~infected + treatment,data=females,type="percent")

```



This graph shows that the FEMALES who got the placebo were more infected than the FEMALES that were vaccinated, but about 50% that got placebos were OK, and about 51% of FEMALES with the vaccine were OK.


#### Numerical Results

```{r}
table1 <- xtabs(~treatment + infected,data=females)
rowPerc(table1)
colPerc(table1)
```



The numerical results reveal that the COVID-19 vaccine significantly decreases the likelihood of contracting the virus compared to the placebo. Specifically, those who receive the vaccine have approximately 46.9% lower odds of getting COVID-19 compared to those who receive the placebo. The relative risk reduction is about 45.5%, indicating that the vaccine reduces the risk of infection by roughly 45.5% relative to the placebo.



#### Inferential Results

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

The chi-square test result of 0.00000001914 and Fisher’s Exact Test result of 0.00000001246 both confirm the statistical significance of the difference in infection rates between the vaccine and placebo groups. The odds ratio of 1.8562 indicates that the likelihood of contracting COVID-19 with the placebo is about 1.86 times higher than with the vaccine. The 95% confidence interval, ranging from 1.49 to 2.32, further supports the effectiveness of the vaccine.




## LGBTQ



$H_0$: The vaccine is not effective in preventing Covid-19 in the LGBTQ community


$H_A$: The vaccine is effective in preventing Covid-19 in the LGBTQ community



## Methods

LGBTQ is CAT
Treatment is CAT
Helper File I used was CAT ~ CAT




```{r}
LGBTQData = filter(FinalData, LGBTQ == "gay")
```



## Descriptive Results
### Graphical Results



```{r}
barchartGC(~infected+treatment,data=LGBTQData)
barchartGC(~infected+treatment,data=LGBTQData, type="percent")
```


This graph shows that the LGBTQ who got the placebo were more infected than the LGBTQ that were vaccinated, but about 45% that got placebos were OK, and about 49% of LGBTQ with the vaccine were OK.



### Numerical Descriptive Results



```{r}
table3 <- xtabs(~infected + treatment, data=LGBTQData)
rowPerc(table3)
colPerc(table3)
```


The numerical results show that the COVID-19 vaccine is effective in lowering the risk of contracting the virus compared to the placebo. Specifically, the odds of contracting COVID-19 with the vaccine are about 24% of those with the placebo, and the risk of infection with the vaccine is approximately 44.8% of that with the placebo. Additionally, the vaccine provides a relative risk reduction of approximately 55.2%, indicating a significant protective effect.

## Infrenital Results



```{r}
chisq.test(table3)
chisqtestGC(table3)
fisher.test(table3)
```



The Chi-squared test yields a p-value of 0.000002266, which is very low. This indicates that the vaccine does significantly reduce the risk of contracting COVID-19 among the LGBTQ community. Additionally, the Fisher's Exact Test provides a p-value of 0.0000007269. The odds of getting COVID-19 with the vaccine are approximately 0.24 times the odds of getting it with the placebo, suggesting that the vaccine considerably lowers the risk compared to the placebo. However, we do not reject the null hypothesis, as individuals in the LGBTQ community can still contract COVID-19 regardless of whether they receive the vaccine or the placebo.



## DrugUsers


$H_0$: The vaccine is not effective in preventing Covid-19 in drug users.


$H_A$: The vaccine is effective in preventing Covid-19 in drug users.



## Methods

DrugUser is CAT
Treatment is CAT
Helper File I used was CAT ~ CAT




```{r}
DruggiesData = filter(FinalData, DrugUser == "yes")
```


## Descriptive Results
### Graphical Results



```{r}
barchartGC(~infected+treatment,data=DruggiesData)
barchartGC(~infected+treatment,data=DruggiesData, type="percent")
```


In this graph it shows that the Drug Users that got the placebo are less likely to get covid19 then those who got the vaccine. But both were about 50% ok.

### Numerical Descriptive Results



```{r}
table4 <- xtabs(~infected + treatment, data=DruggiesData)
rowPerc(table4)
colPerc(table4)
```

The data reveals a 3.28% difference in COVID-19 infection rates between the groups. Specifically, 3.61% of those who received the vaccine contracted COVID-19, compared to 0.33% of those who received the placebo. Among those who remained healthy, there was a difference of only about 2.61%. However, this analysis presents an unusual finding: the vaccine seems to increase the risk of infection compared to the placebo, contrary to expectations. This result could indicate an anomaly in the data or a specific aspect of the study that warrants further investigation. Typically, one would expect a vaccine to reduce, rather than increase, the risk of COVID-19.




## Infrenital Results



```{r}
chisq.test(table4)
chisqtestGC(table4)
fisher.test(table4)
```



The p-value from the Chi-square Test is 0.0000000001414, which is extremely low, suggesting that the vaccine does not effectively prevent COVID-19 among drug users. Similarly, the Fisher's Exact Test yields a p-value of 0.000000000003261, indicating a minor difference in COVID-19 rates among drug users. We do not reject the null hypothesis because drug users can still contract COVID-19 regardless of whether they receive the vaccine or not.